Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia

### 2017-06-30 doc/manuscript/manuscript.Rmd
rerun make
begin pruning introduction

@@

Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
Ascospores infect senescent plant tissues that, when supplemented with cool and humid micro-climate within the plant canopy, leads to mycelial infection of the host plant.
Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen [@abawi1975source].
Diseased bean fields initially show scattered wilted leaves.
Symptoms on individual plants are water-soaked lesions on leaves, pods and stems, which expand over time and show signs of white fungal mycelia.
Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth.

@@

A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
While all three populations were  clonal, the index of association ($I_A$) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S. sclerotiorum* [@ekins2011population].

@@

Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe.
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic].
Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J. Steadman).
Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide.
This information is also helpful in making disease management decisions and developing disease resistance screening methods.

Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
In the case of plant pathogens like *S. sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.

@@

![
Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}

@@

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/Figure3Z.pdf){#community-graph width=50%}

>>>>

<!-- 
ZNK: The introduction needs a lot of work. Here's an outline of what I think
     needs to go in here.
Outline:

 - *S. sclerotiorum* basic biology
 - What have genetic studies and phenotypic studies found
 - What are we missing? A detailed analysis of *S. sclerotiorum* from diverse
   regions that may potentially identify genotypes or MCGs that may be good
   tools for testing breeding efforts.
 - Our hypotheses and specific questions.
 -->

@@

<!-- ZNK: TMI -->
<!-- Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
Ascospores infect senescent plant tissues that, when supplemented with cool and humid micro-climate within the plant canopy, leads to mycelial infection of the host plant.
Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen [@abawi1975source].
Diseased bean fields initially show scattered wilted leaves.
Symptoms on individual plants are water-soaked lesions on leaves, pods and stems, which expand over time and show signs of white fungal mycelia.
Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth. -->

@@

<!-- ZNK: This is more discussion-material -->
<!-- A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
While all three populations were  clonal, the index of association ($I_A$) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S. sclerotiorum* [@ekins2011population].  -->

@@

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
Understanding genetic relationships and reproduction behavior of *S. sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide. 
This information is also helpful in making disease management decisions and developing disease resistance screening methods.

<!-- Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
In the case of plant pathogens like *S. sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.  -->

@@

![
Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
Singletons (46) have been removed for clarity.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
](../../results/figures/publication/Figure1Z.pdf){#barplot width=70%}

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}

@@
